---
title: "Spatial Analysis"
author: "Tirth Gandhi"
date: "2024-07-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Spatial Analysis


```{r data}
data=read.csv("./Data/311_analysis_data.csv")

```


First, let's examine exactly how much of NYC uses a 311 hotline by just plotting all available latitudes and longitudes. 
```{r map}
library(ggplot2)
ggplot(data) +
  geom_point(aes(x=Longitude,y=Latitude),size=0.1)

```
The 311 hotline is used all over NYC and is very popular amongst all NYC residents. Next, we will examine the duration where applicable. 
```{r Map of NYC and duration}
# Load necessary libraries
library(tigris)
library(tmap)
library(sf)
library(dplyr)


# Subset and preprocess the data
subset_data <- data[, c("Latitude", "Longitude", "Duration")]
subset_data$Duration <- abs(subset_data$Duration)
subset_data <- na.omit(subset_data)

# Define the thresholds
threshold1 <- 1440   # 1 day (24 hours)
threshold2 <- 10080  # 1 week (7 days)

# Create a new column for color coding based on the thresholds
subset_data$ColorGroup <- cut(subset_data$Duration,
  breaks = c(-Inf, threshold1, threshold2, Inf),
  labels = c("Within 1 Day", "1 Day to 1 Week", "Beyond 1 Week")
)

# Convert the subsetted data frame to a spatial points data frame
subset_data_sf <- st_as_sf(subset_data, coords = c("Longitude", "Latitude"), crs = 4326)

# List of NYC counties corresponding to the five boroughs
nyc_counties <- c("New York", "Kings", "Bronx", "Queens", "Richmond")

# Get the tract data for all NYC counties
nyc_tracts <- do.call(rbind, lapply(nyc_counties, function(county) {
  tracts(state = 'NY', county = county, cb = TRUE)
}))

# Plot the tracts and points using tmap
tm_shape(nyc_tracts) +
  tm_borders() +
  tm_shape(subset_data_sf) +
  tm_dots(col = "ColorGroup", size = 0.1, alpha = 0.5,
          palette = c("Within 1 Day" = "lightblue", "1 Day to 1 Week" = "orange", "Beyond 1 Week" = "red"), 
          title = "Duration (minutes)")



```
The duration is varied because most complaints take a long time to resolve. Let's examine what complaint types take longer than 24 hours to resolve.

```{r}
filtered_data <- data %>% filter(Duration > 1440 & Duration < 10080)

# Count the occurrences of each Complaint.Type
complaint_counts <- filtered_data %>%
  group_by(Complaint.Type) %>%
  summarise(Count = n()) %>%
  ungroup() %>%
  arrange(desc(Count)) %>%
  slice(1:10)
complaint_counts

```
```{r Proportion of Complaints taking longer than 24 hours and less than 1 week}
sum(data$Duration>1440,na.rm=TRUE)/length(data$Duration)-sum(data$Duration>10080,na.rm=TRUE)/length(data$Duration)

```
```{r Longer than a Week, any overlaps}
filtered_data <- data %>% filter(Duration > 10080)

# Count the occurrences of each Complaint.Type
complaint_counts <- filtered_data %>%
  group_by(Complaint.Type) %>%
  summarise(Count = n()) %>%
  ungroup() %>%
  arrange(desc(Count)) %>%
  slice(1:10)
complaint_counts

```

```{r}
sum(data$Duration>10080,na.rm=TRUE)/length(data$Duration)
```
```{r}
# Filter for cases with ratio_200K_or_more > 0.1
combined_subset <- nrow(data[data$Status == "Closed" & data$Duration < 1440 & data$ratio_200K_or_more > 0.1, ])
# Count rows with Status == "Closed", Duration between 1440 and 10080, and ratio_200K_or_more > 0.1
combined_subset_2 <- nrow(data[data$Status == "Closed" & data$Duration >= 1440 & data$Duration < 10080 & data$ratio_200K_or_more > 0.1, ])
# Count rows with Status == "Closed", Duration > 10080, and ratio_200K_or_more > 0.1
combined_subset_3 <- nrow(data[data$Status == "Closed" & data$Duration > 10080 & data$ratio_200K_or_more > 0.1, ])
# Count rows with Status == "Closed", Duration < 1440, and ratio_200K_or_more < 0.1
combined_subset_4 <- nrow(data[data$Status == "Closed" & data$Duration < 1440 & data$ratio_200K_or_more < 0.1, ])
# Count rows with Status == "Closed", Duration between 1440 and 10080, and ratio_200K_or_more < 0.1
combined_subset_5 <- nrow(data[data$Status == "Closed" & data$Duration >= 1440 & data$Duration < 10080 & data$ratio_200K_or_more < 0.1, ])
# Count rows with Status == "Closed", Duration > 10080, and ratio_200K_or_more < 0.1
combined_subset_6 <- nrow(data[data$Status == "Closed" & data$Duration > 10080 & data$ratio_200K_or_more < 0.1, ])
combined_subset/combined_subset_4
combined_subset_2/combined_subset_5
combined_subset_3/combined_subset_6

```
```{r}
mean_duration_1 <- mean(data[data$Status == "Closed" & data$Duration < 1440 & data$ratio_200K_or_more > 0.1, "Duration"], na.rm = TRUE)
mean_duration_2 <- mean(data[data$Status == "Closed" & data$Duration >= 1440 & data$Duration < 10080 & data$ratio_200K_or_more > 0.1, "Duration"], na.rm = TRUE)
mean_duration_3 <- mean(data[data$Status == "Closed" & data$Duration > 10080 & data$ratio_200K_or_more > 0.1, "Duration"], na.rm = TRUE)
mean_duration_4 <- mean(data[data$Status == "Closed" & data$Duration < 1440 & data$ratio_200K_or_more < 0.1, "Duration"], na.rm = TRUE)
mean_duration_5 <- mean(data[data$Status == "Closed" & data$Duration >= 1440 & data$Duration < 10080 & data$ratio_200K_or_more < 0.1, "Duration"], na.rm = TRUE)
mean_duration_6 <- mean(data[data$Status == "Closed" & data$Duration > 10080 & data$ratio_200K_or_more < 0.1, "Duration"], na.rm = TRUE)
mean_duration_1/mean_duration_4
mean_duration_2/mean_duration_5
mean_duration_3/mean_duration_6

```
Let's shift our focus from duration to which locations are causing the most complaints by plotting the number of times a particular location appears. 
```{r }
library(dplyr)
library(sf)
library(tmap)

# Calculate location counts and apply threshold
location_counts <- data %>%
  group_by(Latitude, Longitude) %>%
  summarise(Count = n()) %>%
  ungroup()

location_counts <- na.omit(location_counts)

# Define threshold and add ColorGroup and SizeGroup
threshold <- 100
location_counts <- location_counts %>%
  mutate(
    ColorGroup = ifelse(Count > threshold, "Above 100 Complaints", "Below 100 Complaints"),
    SizeGroup = ifelse(Count > threshold, 2, 0.1)
  )

# Convert to sf object
location_counts_sf <- st_as_sf(location_counts, coords = c("Longitude", "Latitude"), crs = 4326)

# Define color palette with fixed colors
color_palette <- c("Below 100 Complaints" = "lightblue", 
                   "Above 100 Complaints" = "darkred")

# Plot the data using tmap
tm_shape(location_counts_sf) +
  tm_dots(size = "SizeGroup", col = "ColorGroup", palette = color_palette, title = "Complaints per Location") +
  tm_layout(title = "Complaint by Location Counts", title.size = 1)



```
The most problematic areas are near bodies of water where more people are gathered which is consistent with our prior findings and near public spaces like parks, even the few darkred spots in-land are near public spaces or business districts when you cross-verify with Google Maps. Below is a direct comparison about other related factors with respect to Complaints per location.  







```{r Direct Comparison}
problematic_locations <- location_counts %>%
  filter(ColorGroup == "Above 100 Complaints")

non_problematic_locations <- location_counts %>%
  filter(ColorGroup == "Below 100 Complaints")

# Merge with original data
merged_problematic <- data %>%
  right_join(problematic_locations, by = c("Latitude", "Longitude"))

merged_non_problematic <- data %>%
  right_join(non_problematic_locations, by = c("Latitude", "Longitude"))

# Select relevant columns and calculate summary statistics
calculate_summary <- function(df) {
  df %>%
    summarise(
      ratio_200K_or_more_avg = mean(ratio_200K_or_more, na.rm = TRUE),
      ratio_200K_or_more_median = median(ratio_200K_or_more, na.rm = TRUE),
      ratio_finished_higher_degree_avg = mean(ratio_finished_higher_degree, na.rm = TRUE),
      ratio_finished_higher_degree_median = median(ratio_finished_higher_degree, na.rm = TRUE),
      Median_age_years_avg = mean(Median_age_years, na.rm = TRUE),
      Median_age_years_median = median(Median_age_years, na.rm = TRUE),
      Male_percentage_avg = mean(Male_percentage, na.rm = TRUE),
      Male_percentage_median = median(Male_percentage, na.rm = TRUE),
      Female_percentage_avg = mean(Female_percentage, na.rm = TRUE),
      Female_percentage_median = median(Female_percentage, na.rm = TRUE)
    )
}

summary_problematic <- calculate_summary(merged_problematic)
summary_non_problematic <- calculate_summary(merged_non_problematic)

# Combine the summaries into a single dataframe
comparison_summary <- bind_rows(
  problematic = summary_problematic,
  non_problematic = summary_non_problematic,
  .id = "Group"
)

# Display the resulting dataframe
print(comparison_summary)


```



